Ripple Consensus Review

### 2015-04-26 paper.tex
More fleshed out attacks

\subsection{Choosing the Unique Node List}

@@

    \item \textbf{Scope:} Who is affected by the attack? Is this a targetted attack,
          mainly affecting a one or two targets? An attack with broad impact? Or
          an attack with global impact on all Ripple users?

@@

big picture with attacks is there is not a minimum cost to them unlike PoW

evaluate attacks for denial-of-service and theft potential; evaluate minimum
cost to carry out attack; possible minimum manpower to carry out attack

Type of attack: DoS and/or theft

Minimum cost: $0, $1k, $10k, $100k

Scope: targetted, broad, global
Duration: hours, days, weeks, indefinite - How long will it take to fix the bug?

we don't attempt to put a dollar figure on harm done, as depends enormously on
how people use ripple

@@

@@

DoS: Cost \$0, Scope: global, Duration: hours, Prob: high
Theft: Cost \$1k, Scope: varies, Duration: hours, Prob: high

not as synergistic with other attacks; obvious flaw that will be fixed;
time-limited attack

@@

DoS: Cost depends, Scope: broad, Duration: depends

@@

\subsection{UNL: Jurisdictional}

DoS/Theft: Cost \$100k, Scope: varies, Duration: indefinite

centralization attack - legal attack against centralized UNL. note
jurisdictional issues with delegating consensus to UNL

biggest issue is how can ripple compete against systems without the global
consensus req? if I'm a bank in russia, why use ripple when it opens me to UNL
issues?

@@

DoS/Theft: Cost \$0, Scope: varies, Duration: varies

mention unsigned code here

\subsection{UNL: key theft}

prep: Cost \$0, Scope: global, Duration: N/A

Gain control of UNL signing keys through theft/compromise.

\subsection{Validator Fear, Uncertainty, and Doubt}

make people afraid to run validating nodes

illegal data

\subsection{History simulation}

Theft: Cost \$0, Scope: targetted, Duration: weeks

from proof-of-stake terminology; simulated history is costless as no work
expended.

use backdoored software as example of how this can happen; also can happen with
jurisdictional attacks

if rippled sees two versions of history, what happens?

rewrite history attack - note lack of cost to rewrite because no PoW; single
implementation so likely we can hack all nodes at once

proof-of-work has a hard bound on rewrite attacks, because can't break laws of
physics

can talk about how cost to do a simulation attack is PoW targets; talk about slasher rule

>>>>

\subsection{Choosing the Unique Node List}
\label{choosing-the-unl}

@@

    \item \textbf{Scope:} Who is affected by the attack? Is this a targeted attack,
          mainly affecting a one or two targets? An attack with broad impact? Or
          an attack with global impact on all Ripple users?

@@

@@

\begin{center}
    \begin{tabular}{c|c|c|c|c}
        Type & Cost & Scope & Duration & Probability \\ \hline
        DoS   & $\$0$ & Global & Hours & Very High \\
        Theft & $\$1\text{k}$ & Targeted & Hours & Medium
    \end{tabular}
\end{center}

@@

@@

\begin{center}
    \begin{tabular}{c|c|c|c|c}
        Type & Cost & Scope & Duration & Probability \\ \hline
        DoS & $>\$100\text{k}$ & Broad & Varies & Low
    \end{tabular}
\end{center}

@@

\subsection{Coercion of validators}

\begin{center}
    \begin{tabular}{c|c|c|c|c}
        Type & Cost & Scope & Duration & Probability \\ \hline
        DoS & $>\$100\text{k}$ & Targeted & Indefinite & High \\
        DoS & $>\$1\text{M}$ & Global & Indefinite & Medium \\
        Theft & $>\$100\text{k}$ & Targeted & Indefinite & Medium \\
        Theft & $>\$1\text{M}$ & Global & Indefinite & Low
    \end{tabular}
\end{center}

Validators are coerced through poltical/legal/criminal means into attacking
some or all Ripple users by preventing valid transactions from entering the
ledger (DoS) and/or by allowing invalid transactions to enter the ledger.
(theft) The cost estimates above are very rough, and assuming the attack is via
legal means; the cost may potentially be as low as zero in some scenarios like
blackmail. In preparation for this attack the attackers may also attempt to
reduce the number of validators by with, for instance, a transaction flood
attack.

    \item The default UNL list and incentives against operating or using
        non-official validators discussed in Section \ref{choosing-the-unl}
        significantly increase the effectiveness of a coercion attack as nearly
        all Ripple users will be affected. Equally, in the event of a coercion
        attack there are still significant downsides to adopting a different
        UNL.

    \item Similarly this attack is synergistic with transaction flooding and
        other attacks on the scalability of the network, reducing the number of
        validators that need to be coerced.

    \item The rippled node implementation does not check full history. This
        means if validators sign an invalid ledger, the transactions that make
        it invalid can be hidden from Ripple users. Equally, even with
        detection, the fact that Ripple users can be asked to simply
        restart/reinstall their nodes to get past the invalid history may
        significantly reduce the social barriers to creating an invalid ledger.

    \item The Ripple protocol is constructed such that it is possible to create
        many, although not all, types of fraud
        proofs\cite{irc-btcdev-gmaxwell-fraud-proof, btctalk-pkt-fraud-proofs}
        required to prove an invalid ledger. While the software does not
        currently implement fraud proofs, this is functionality that could be
        added in the future without a change to the consensus. If widely used
        fraud proofs could prevent some types of coercion attacks that force
        validators to create invalid ledgers by making it likely for the result
        of the attack to be DoS rather than profitable theft. In short, the
        attack would simply shut the Ripple network down for all users rather
        than result in a succesful theft.

    \item Some people would argue that the recently announced
        plans\cite{cointelegraph-ripple-aml} to require verified user
        identification to use the Ripple network in the near future are an
        example of a succesful coercion attack. It can be argued that Ripple
        Labs is increasing the risk of coercion attacks by simultaneously
        adopting a ``regulator friendly'' position and promoting the use of
        Ripple across many different jurisdictions.

@@

\begin{center}
    \begin{tabular}{c|c|c|c|c}
        Type & Cost & Scope & Duration & Probability \\ \hline
        Varies & $\$0$ & Broad & Weeks & High \\
    \end{tabular}
\end{center}

The attacker inserts a backdoor into a widely used Ripple protocol
implementation. (e.g. the rippled node software) The attacker may be an
internal Ripple Labs employee, an external member of the public submitting a
pull-request, or they may compromise the software actually downloaded by
end-users by compromising binary or sourcecode hosting. (e.g. GitHub)

    \item All relevant software is open source, allowing for third-party
        review. Ripple Labs also has a bug bounty program, which can
        incentivise that review.

    \item Ripple Labs appears to use GitHub - a third party -
        as their master sourcecode repository, with development happening
        through pull-requests on GitHub. While this does place trust in a
        third-party, PGP signing sourcecode during development can mitigate
        this risk. However there appears to be no PGP signatures on any git
        commits, giving Ripple Labs little ability to audit who wrote what
        code.

    \item \textbf{Currently Ripple Labs does not provide a secure way to
        download any of their software.} While binaries are available, the
        mechanism to download them is a Ubuntu package repository over insecure
        HTTP; while source code is available through git, neither commits nor
        release tags are signed in anyway. \textbf{This is a serious ommission
        that has lead to significant monetary losses in the past.} Ripple Labs
        should be following industry best-practice by signing git commits and
        tags\cite{wladimir-git-pgp} as well as PGP signing their Ubuntu
        packages.

\subsection{Theft of validator secret keys}

\begin{center}
    \begin{tabular}{c|c|c|c|c}
        Type & Cost & Scope & Duration & Probability \\ \hline
        Precursor & $\$0$ & N/A & N/A & Unknown \\
    \end{tabular}
\end{center}

The attacker steals illegitimate control of the secret keys used by validators,
particularly the validators in the default UNL. The theft may happen via an
exploit, backdoor, compromised hosting company, etc. While not useful for the
attacker in of itself, the stolen keys can be used for further atttacks such as
DoS or a simulated ledger attack.

A closely related attack is to compromise the Ripple website, and trick people
into downloading a default UNL list consisting of only attacker controlled
nodes.

    \item Due to the high centralization of UNL validators, a relatively small
        number of validators need to be compromised for a succesful attack.

    \item Most validators are running nearly the same, if not the exact same,
        software, making it likely for backdoor/exploit attack to succesfully
        exfiltrate the secret keys. Due to the high consensus-critical
        complexity of the Ripple protocol and lack of a consensus library
        alternate validator implementations are currently infeasible, making it
        difficult to change this situation.

    \item The rippled implementation doesn't yet appear to support separating
        the validation/consensus and signing functionality into different
        machines, or even different user accounts. (possibly Ripple Labs has
        internal-use-only improvements)

    \item Details about the operational security practices of the validators on
        the default UNL is unavailable, making it difficult to assess risk.

    \item The Ripple P2P protocol does make it possible to operate validators
        without the nodes themselves accepting public traffic; the signatures
        produced by the validating nodes are distributed via an anonymizing
        flood-fill network. This setup is recommended by Ripple Labs
        documentation and makes remote attacks significantly more difficult.

\subsection{Simulated ledger}

\begin{center}
    \begin{tabular}{c|c|c|c|c}
        Type & Cost & Scope & Duration & Probability \\ \hline
        Theft & $>\$1\text{k}$ & Targeted & Days & N/A
    \end{tabular}
\end{center}

An attacker with the ability to create signatures from nodes on the UNL can
create an entirely simulated ledger in parallel to the ``official'' Ripple
ledger, containing any transaction the attacker wishes. In conjunction with
sybil attacking the target the attacker can then perform double-spend attacks.
Sybil attacks are relatively cheap, requiring just the aquisition of a large
number of IP addresses, e.g. via cloud hosting and/or botnet rental. In some
situations sybil attacks can even have zero cost, such as the internal IT staff
within the target with access to network infrastructure, the target's ISP, or
government actors with MITM capabilities such as the
NSA.\cite{schneier-nsa-mitm}

    \item Because the Ripple consensus is determined by signatures creating a
        simulated ledger is a zero-cost attack, the ``costless
        simulation''\cite[4.2]{apoelstra-on-stake} problem. Proof-of-work
        ledgers in comparison have very well-defined costs that discourage
        attacks and incentivise detection.\footnote{On the Bitcoin network even
            if the attacker can steal the hashing power needed to simulate
        history for free the miners involved lose money from the attack and
    have strong monetary incentives to detect and stop it.}

    \item There are no direct incentives built into the Ripple protocol to
        discourage validators from allowing their private keys to sign
        simulated ledgers. By comparison the Slasher proof-of-stake algorithm
        attempts to disincentivise simulated history by destroying valuable
        bonds owned up by participants in the consensus
        process.\cite{ethereum-blog-slasher}

    \item The lack of incentive may invite coercion attacks, as the validators
        being coerced into signing a simulated ledger have no direct incentive
        not too. For instance a court could order a validator to assist in a
        simulation attack on a target as a way of recovering funds held by the
        target.

    \item Simulation attacks are made significantly easier by the fact that the
        Ripple consensus protocol will remove validators from the UNL if they
        appear to have failed.\cite[3.4.2]{ripple-consensus-paper} With the
        target sybil attacked the attacker can prevent the target from learning
        that the other validators have \emph{not} failed. When those validators
        are removed from the UNL due to the simulated failure the attacker does
        not need to compromise/coerce as many validators as before. Equally, if
        the target is using network split detection, the attacker can make the
        non-compromised validators \emph{actually} fail as preparation for the
        attack. Again, the lack of incentive to run a validator is problematic.
